Promise and peril: Agnostic identification methods\\ for detecting differential item functioning

### 2020-10-14 paper/paper.tex
apparently i did a bunch of stuff and never committed it

\title{Promise and peril: Agnostic identification methods\\
for detecting differential item functioning}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{Ben Stenhaug, Ben Domingue, and Mike Frank\\
Stanford University}
\author{}
\date{\vspace{-2.5em}}

\begin{document}
\maketitle
\begin{abstract}
It is well known that likelihood ratio tests (LRT) are effective at detecting differential item functioning (DIF) in item response models. However, to use a LRT (or any other method), we require an identifying assumption to disentangle differences in group ability from potential DIF. We use the term ``agnostic identification'' (AI) to describe the process of finding such identifying assumptions without a priori knowledge of relative group ability or items that are DIF-free. We first summarize existing AI methods and propose a variety of new methods. We then conduct a simulation study---which we argue is more realistic than most DIF simulation studies in the literature---and find that one of the proposed new AI methods, all-others-as-anchors-one-at-a-time (AOAA-OAT), significantly outperforms current methods. We also suggest a new approach, the equal means, multiple imputation logit graph (EM-MILG), which clearly presents all information about possible DIF, including sampling variability in item parameters, to the analyst. \clearpage
\end{abstract}

@@

Despite this, the usual setup of DIF simulation studies frames DIF as a property of the item. For example, Kopf, Zeileis, and Strobl (\protect\hyperlink{ref-kopf2015framework}{2015}) simulate students as belonging to either a reference or focal group. They fix the item easinesses for the reference group, \({b_j}^{\text{ref}}\), to values that they obtained from a previous study. They set item easinesses for the focal group to \({b_j}^{\text{foc}} = {b_j}^{\text{ref}}\) for items without DIF and \({b_j}^{\text{foc}} = {b_j}^{\text{ref}} - 0.6\) for items with DIF, where 0.6 is the magnitude of DIF in logits. They then simulate student ability \({\theta_i}^{\text{ref}} \sim N(0,1)\) for students in the reference group and \({\theta_i}^{\text{foc}} \sim N(-1,1)\) for students in the focal group. Finally, they generate item responses according to the Rasch model, which specifies that the probability of student \(i\) responding correctly to item \(j\) is
\begin{align}
\text{Pr}(y_{ij} = 1 | \theta_i, b_j) = \sigma(\theta_i + b_j) \label{zopf}
\end{align}
where \(\sigma(x) = \frac{e^x}{e^x + 1}\) is the standard logistic function.

For every DIF simulation study framed in terms of item parameters that vary across groups, there is a mathematically equivalent setup in which students' abilities are multidimensional. According to the two-dimensional compensatory 2PL model (Thissen and Steinberg \protect\hyperlink{ref-thissen1986taxonomy}{1986}), the probability that student \(i\) responds correctly to item \(j\) is
\begin{align}
\text{Pr}(y_{ij} = 1 | \theta_i, \eta_i, a_{\theta j}, a_{\eta j}, b_j) = \sigma(a_{\theta j}\theta_i + a_{\eta j}\eta_i + b_j).
\end{align}
where \(\theta\) is target ability, \(\eta\) is nuisance ability, \(a_{\theta j}\) is an item's slope on target ability, and \(a_{\eta j}\) is an item's slope on nuisance ability.

For example, to translate the Kopf, Zeileis, and Strobl (\protect\hyperlink{ref-kopf2015framework}{2015}) simulation from the DIF-as-item-property to the DIF-as-student-property view, \(b_j\) is set to \({b_j}^{\text{ref}}\), \(\theta\) stays the same, \(a_{\theta j} = 1\) for all items (consistent with the Rasch model), and \(\eta\) takes one of two values: \({\eta_i}^{\text{ref}} = 0\) for students in the reference group and \({\eta_i}^{\text{foc}} = -1\) for students in the focal group. For items with DIF, \(a_{\eta j} = 0.6\), and for items without DIF, \(a_{\eta j} = 0\) Again, 0.6 is the magnitude of DIF in logits.\footnote{The reader may notice that this is just one possible specification. Any specification that meets the following criteria has the same effect: First, \(a_{\eta j}\eta_i\) must be 0.6 lower for the focal group on items with DIF. Second, \(a_{\eta j}\eta_i\) must be equal across groups for items without DIF.} For a choice of student (from the focal or reference group) and item (with DIF or without), this model produces identical probabilities as Eqn \ref{zopf}. This translation between views makes explicit that nearly all DIF simulation studies have, perhaps suboptimally, examined the unrealistic scenario in which there is no variation in the nuisance ability for students in the same group.

If we insist on describing simulation conditions from the DIF-as-student-property view, one might wonder the following: Why not fit a multidimensional item response model which describes the data fully instead of looking for bias in a lower dimensional model? Camilli (\protect\hyperlink{ref-camilli1992conceptual}{1992}) tested this idea with the goal of a \enquote{more satisfying description of the secondary abilities} {[}p.~144{]}. He found that the rotational indeterminacy of item response models is challenging to overcome and concluded that \enquote{a priori knowledge of the true factor structure} is necessary {[}p.~144{]}. It's hard to imagine how one would have such knowledge. Therefore, the best approach, which the DIF literature has nearly unanimously taken, is to fit unidimensional item response models and then look for bias manifesting in the item parameters. This approach depends upon a crucial identifying assumption; this assumption is the target of our investigation.

@@

Psychometricians have long been in search of the perfect method for detecting differential functioning of \emph{individual} items. The IRT-based likelihood ratio test (LRT) is effective at detecting DIF (Meade and Wright \protect\hyperlink{ref-meade2012solving}{2012}). Given both the efficacy or the LRT approach and our preference for methods that existin within the IRT-framework, we do not discuss methods like the Mantel-Haenszel procedure (Holland and Thayer \protect\hyperlink{ref-holland1986differential}{1986}), which has been shown to perform no better---and in some cases worse---than IRT-based methods (Swaminathan and Rogers \protect\hyperlink{ref-swaminathan1990detecting}{1990})). Within the IRT-framework, we---as is common---estimate models using marginal maximum likelihood estimation (Bock and Aitkin \protect\hyperlink{ref-bock1981marginal}{1981}) and identify the scale by setting \(\mu^\text{ref} = 0\).

Use of LRT requires an identifying assumption. The assumption is needed so as to link groups. In the typical setting, no a prior assumptions---about either relative group ability or which items might have DIF---can be made. Thus, we're interested in \emph{agnostic identification} methods (hereafter referred to as \enquote{AI methods}). The fundamental challenge is to disentangle estimation of \(\hat\mu^\text{foc} - \mu^\text{ref} = \hat\mu^\text{foc} - 0 = \hat\mu^\text{foc}\), the difference in mean abilities, from the estimation of \(\hat d = \hat{b_j}^{\text{foc}} - \hat{b_j}^{\text{ref}}\), the difference in item easiness. There are two types of AI methods: The most common, the use of anchor items, selects a group of items that are assumed to be DIF-free. These anchor items identify the model, thereby allowing for the estimation of \(\hat\mu^\text{foc}\), and the remaining items can be tested for DIF using an LRT. The other type, the use of an anchor point (Strobl et al. \protect\hyperlink{ref-strobl2018anchor}{2018}), directly sets the difference in mean abilities to some value, \(\mu^{\star\text{foc}}\). This anchor point identifies the model, and all items can be tested for DIF using an LRT.

In this section, we summarize existing AI methods and propose a variety of extensions. For simplicity, we focus on the Rasch model to isolate the fundamental issues in \enquote{the AI problem}. Further, so as to offer a coherent framework, we sometimes edit names of existing methods.\footnote{We recognize that others have done the same (e.g., Kopf, Zeileis, and Strobl \protect\hyperlink{ref-kopf2015framework}{2015}), and that we risk contributing to a proliferation of names.} We use a (simulated) demonstration dataset to illustrate key features of the different methods.\footnote{This data is used solely for illustrative purposes; we describe an extensive simulation study later in the paper.} The data consists of 10,000 reference group students and 10,000 focal group students taking an eight-item test. Our simulation is similar to our multidimensional alternative to the Kopf, Zeileis, and Strobl (\protect\hyperlink{ref-kopf2015framework}{2015}) approach discussed above. Target ability is simulated as \({\theta_i}^{\text{ref}} \sim N(\mu^\text{ref} = 0,1)\) for students in the reference group and \({\theta_i}^{\text{foc}} \sim N(\mu^\text{foc} = -1,1)\) for students in the focal group. Nuisance ability is set to \({\eta_i}^{\text{ref}} = 0\) and \({\eta_i}^{\text{foc}} = -1\). The slope on target ability is set to \(a_{1j} = 1\) for all items. The slope on nuisance ability is set to \(a_{2j} = 0.5\) for the last three items (the items with DIF) and \(a_{2j} = 0\) otherwise. We can also, of course, describe these conditions from the DIF-as-item-property view where there is no nuisance ability. Instead, \({b_j}^{\text{foc}} = {b_j}^{\text{ref}}\) for the first five items and \({b_j}^{\text{foc}} = {b_j}^{\text{ref}} - 0.5\) for the last three items. To be sure, this data is used only for demonstrative purposes. We compare the performance of the AI methods in a more realistic scenario in the \protect\hyperlink{simstudy}{the simulation study}.

The AI methods described in this paper are summarized in Table \ref{table:allmethods}. In addition the results for each method on the demonstration data are shown in a MILG in Figure \ref{fig:allmethodsfig}.

@@

AI methods typically work like a black box. The analyst puts their item response data in and the black box outputs an identification assumption to be used when fitting subsequent models.

While we will document the performance of these methods momentarily, we first suggest an approach focused on putting analysts in a better position to understand the scale of the problem in their data. We propose the \enquote{equal means, multiple imputation logit graph} (EM-MILG), which presents information about potential DIF to the analyst. EM-MILG begins by fitting a unidimensional Rasch model to the data that is identified by arbitrarily setting \(\mu^\text{foc} = 0\). Given that we also assume \(\mu^\text{ref} = 0\), this is equivalent to the assumption that the groups have equal mean ability. As a result, all differences in performance---either from group ability differences or DIF---manifest in the item easiness difference parameter \(\tilde{d_j}\). The tilde (e.g., \(\tilde{d_j}\)) is used to denote parameters estimated with \(\mu^\text{foc}\) set to \(0\)---as compared to the hat (e.g., \(\hat{d_j}\)), which is used for parameter estimates from a properly identified model.

@@

Figure \ref{fig:emmilg} shows the EM-MILG for our demonstration data. We emphasize that the EM-MILG contains all possible information about the difference in group performance. The analyst might assume that---because there are five items where the reference group outperforms the focal group by approximately 1 logit and only three items where the difference is 1.5 logits---items 1-5 are unbiased. This might seem to be a strong assumption, but all AI methods assume---in one way or another---that, if there is DIF, its a minority of items that are responsible. These unbiased items are known as anchor items; setting anchor items is the most common identification assumption. After anchor items have been selected, the model is refit where the identifying assumption that \(\mu^\text{foc} = 0\) is dropped and replaced with the identifying assumption that the anchor items are DIF-free (i.e., \(d_j = 0\) for anchor items). Of course, the demonstration data is designed such that there are obvious groups; this will not necessarily typically be the case.

@@

We described a variety of AI methods and their corresponding acronymns. Some of these methods, such as AOAA and EMC, select anchor items, while others, such as MAXGI and MINBC, select an anchor point. Table \ref{table:allmethods} summarizes all of these methods. Figure \ref{fig:allmethodsfig} shows the results for each method on the demonstration data in a MILG.

@@

\begin{verbatim}
## Picking joint bandwidth of 0.00587
\end{verbatim}

\begin{verbatim}
## Picking joint bandwidth of 0.00591
\end{verbatim}

\begin{verbatim}
## Picking joint bandwidth of 0.00546
\end{verbatim}

\begin{verbatim}
## Picking joint bandwidth of 0.0055
\end{verbatim}

@@

\caption{MILG for each of the AI methods applied to the demonstration data. AOAA-OAT and EMC are identified by anchor items (represented by points), MAXGI and MINBC are identified by setting $\mu^{\text{foc}}$ set to the anchor point $\mu^{\star\text{foc}}$, AOAA and AOAA-AS are not included as they did not find an identifying assumption.}\label{fig:allmethodsfig}
\end{figure}

@@

AOAA-OAT exhibiting superior performance is an important finding given that it is, to our knowledge, not used currently. We advocate for its widespread use. One drawback of the AOAA-OAT method is that it is computationally expensive. For example, finding three items containing DIF on a 12-item test requires fitting 46 item response models, and that number grows as either the test length or the number of items testing positive for DIF grows. To increase AOAA-OAT's use, we recommend its implementation (perhaps as the default) in popular IRT software such as the mirt R package.

@@

Moreover, future work should test these methods' performance under a greater variety of data-generating conditions. For example, changing the compensatory nature of the data generating model or adding additional nuisance ability dimensions. Finally, our work focused on the Rasch model, and it is of great interest to consider how these methods extend and perform when the goal is to detect and correct for DIF when fitting other types of item response models.

@@

@@

@@

@@

@@

>>>>

\title{Promise and peril: Agnostic identification methods\\
for detecting differential item functioning}
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{Ben Stenhaug, Mike Frank, and Benjamin W. Domingue\\
Stanford University}
\author{}
\date{\vspace{-2.5em}}

\begin{document}
\maketitle
\begin{abstract}
It is well known that likelihood ratio tests (LRT) are effective at detecting differential item functioning (DIF) in item response models. However, to use a LRT (or any other method), we require an identifying assumption to disentangle differences in group ability from potential DIF. We use the term ``agnostic identification'' (AI) to describe the process of finding such identifying assumptions without a priori knowledge of relative group ability or items that are DIF-free. We first summarize existing AI methods and propose a variety of new methods. We then conduct a simulation study---which we argue is more realistic than most DIF simulation studies in the literature---and find that one of the proposed new AI methods, all-others-as-anchors-one-at-a-time (AOAA-OAT), significantly outperforms current methods. We also suggest a new approach, the equal means, multiple imputation logit graph (EM-MILG), which clearly presents all information about possible DIF, including sampling variability in item parameters, to the analyst. Given the central importance of AI techniques to subsequent examinations of item-level bias, we believe these constitute important advances for the field. \clearpage
\end{abstract}

@@

Despite this, the usual setup of DIF simulation studies frames DIF as a property of the item. For example, Kopf, Zeileis, and Strobl (\protect\hyperlink{ref-kopf2015framework}{2015}) simulate students as belonging to either a reference or focal group. They fix the item easinesses for the reference group, \({b_j}^{\text{ref}}\), to values that they obtained from a previous study (Wang, Shih, and Sun \protect\hyperlink{ref-wang2012dif}{2012}). They set item easinesses for the focal group to \({b_j}^{\text{foc}} = {b_j}^{\text{ref}}\) for items without DIF and \({b_j}^{\text{foc}} = {b_j}^{\text{ref}} - 0.6\) for items with DIF, where 0.6 is the magnitude of DIF in logits. They then simulate student ability \({\theta_i}^{\text{ref}} \sim N(0,1)\) for students in the reference group and \({\theta_i}^{\text{foc}} \sim N(-1,1)\) for students in the focal group. Finally, they generate item responses according to the Rasch model, which specifies that the probability of student \(i\) responding correctly to item \(j\) is
\begin{align}
\text{Pr}(y_{ij} = 1 | \theta_i, b_j) = \sigma(\theta_i + b_j) \label{zopf}
\end{align}
where \(\sigma(x) = \frac{e^x}{e^x + 1}\) is the standard logistic function.

For every DIF simulation study framed in terms of item parameters that vary across groups, there is a mathematically equivalent setup in which students' abilities are multidimensional. The two-dimensional compensatory 2PL model (Thissen and Steinberg \protect\hyperlink{ref-thissen1986taxonomy}{1986}) asserts that the probability that student \(i\) responds correctly to item \(j\) is
\begin{align}
\text{Pr}(y_{ij} = 1 | \theta_i, \eta_i, a_{\theta j}, a_{\eta j}, b_j) = \sigma(a_{\theta j}\theta_i + a_{\eta j}\eta_i + b_j).
\end{align}
where \(\theta\) is target ability, \(\eta\) is nuisance ability, \(a_{\theta j}\) is an item's slope on target ability, and \(a_{\eta j}\) is an item's slope on nuisance ability.

We can use this model to translate the Kopf, Zeileis, and Strobl (\protect\hyperlink{ref-kopf2015framework}{2015}) simulation from the DIF-as-item-property to the DIF-as-student-property view; \(b_j\) is set to \({b_j}^{\text{ref}}\), \(\theta\) stays the same, \(a_{\theta j} = 1\) for all items (consistent with the Rasch model), and \(\eta\) takes one of two values: \({\eta_i}^{\text{ref}} = 0\) for students in the reference group and \({\eta_i}^{\text{foc}} = -1\) for students in the focal group. For items with DIF, \(a_{\eta j} = 0.6\), and for items without DIF, \(a_{\eta j} = 0\). Again, 0.6 is the magnitude of DIF in logits.\footnote{The reader may notice that this is just one possible specification. Any specification that meets the following criteria has the same effect: First, \(a_{\eta j}\eta_i\) must be 0.6 lower for the focal group on items with DIF. Second, \(a_{\eta j}\eta_i\) must be equal across groups for items without DIF.} For a choice of student (from the focal or reference group) and item (with DIF or without), this model produces identical probabilities as Eqn \ref{zopf}.

However, this translation between views reveals an important point. As a side effect of this approach, simulation studies have tended to focus on the unrealistic scenario in which there is no variation in the nuisance ability for students in the same group. Much as students vary in the target ability---a belief in such variation is central to the measurement enterprise---we suspect they also vary on additional ability dimensions. Indeed, previous work has shown that the best fitting model for item response data is often multidimensional where each dimension has significant variation (e.g., Furlow, Raiford Ross, and Gagné \protect\hyperlink{ref-furlow2009impact}{2009}). These dimensions are typically positively correlated, and the greater that correlation, the better unidimensional models tend to perform (Kirisci, Hsu, and Yu \protect\hyperlink{ref-kirisci2001robustness}{2001}). Explicitly accounting for this variation---which we do in our simulation study---should yield results more indicative of how DIF methods will perform in practice.

If we insist on describing simulation conditions from the DIF-as-student-property view, one might wonder the following: Why not fit a multidimensional item response model which describes the data fully instead of looking for bias in a lower dimensional model? Camilli (\protect\hyperlink{ref-camilli1992conceptual}{1992}) tested this idea with the goal of a \enquote{more satisfying description of the secondary abilities} {[}p.~144{]}. He found that the rotational indeterminacy of item response models is challenging to overcome and concluded that \enquote{a priori knowledge of the true factor structure} is necessary {[}p.~144{]}. It's hard to imagine how one would have such knowledge. Therefore, the best approach, which the DIF literature has nearly unanimously taken, is to fit unidimensional item response models and then look for bias manifesting in the item parameters. This approach depends upon a crucial identifying assumption, which we discuss at length below, is the target of our investigation.

@@

Psychometricians have long been in search of the perfect method for detecting differential functioning of \emph{individual} items. The IRT-based likelihood ratio test (LRT) is effective at detecting DIF (Meade and Wright \protect\hyperlink{ref-meade2012solving}{2012}). Given both the efficacy of the LRT approach and our preference for methods that exist within the IRT-framework, we do not discuss alternative methods like the Mantel-Haenszel procedure (Holland and Thayer \protect\hyperlink{ref-holland1986differential}{1986}), which has been shown to perform no better---and in some cases worse---than IRT-based methods (Swaminathan and Rogers \protect\hyperlink{ref-swaminathan1990detecting}{1990})). Within the IRT-framework, we---as is common---estimate models using marginal maximum likelihood estimation (Bock and Aitkin \protect\hyperlink{ref-bock1981marginal}{1981}) and identify the scale by setting \(\mu^\text{ref} = 0\). Both of these choices are---for our purposes---completely inconsequential.

The use of LRT requires an identifying assumption. The assumption is needed so as to link groups. There are two possibilities. One possibility is to utilize information about the distribution of the focal group; in particular, knowledge that the focal group and reference group have the same mean ability would be valuable. Alternatively, we can use information from a set of unbiased items to derive such information. In the equating literature, the former is a \enquote{non-common item random groups} design and the latter is a \enquote{common-item nonequivalent groups} design (Cook and Paterson \protect\hyperlink{ref-cook1987problems}{1987}; Topczewski et al. \protect\hyperlink{ref-topczewski2013comparison}{2013}).

However, in the typical setting, no a prior assumptions---about either relative group ability or which items might have DIF---can be made. Thus, we're interested in \emph{agnostic identification} methods (hereafter referred to as \enquote{AI methods}). The fundamental challenge is to disentangle estimation of \(\hat\mu^\text{foc} - \mu^\text{ref} = \hat\mu^\text{foc} - 0 = \hat\mu^\text{foc}\), the difference in mean abilities, from the estimation of \(\hat d = \hat{b_j}^{\text{foc}} - \hat{b_j}^{\text{ref}}\), the difference in item easiness. There are two types of AI methods: The most common, the use of anchor items, selects a group of items that are assumed to be DIF-free. These anchor items identify the model, thereby allowing for the estimation of \(\hat\mu^\text{foc}\), and the remaining items can be tested for DIF using an LRT. The other type, the use of an anchor point (Strobl et al. \protect\hyperlink{ref-strobl2018anchor}{2018}), directly sets the difference in mean abilities to some value, \(\mu^{\star\text{foc}}\). This anchor point identifies the model, and all items can be tested for DIF using an LRT.

In this section, we summarize existing AI methods and propose a variety of extensions. For simplicity, we focus on the Rasch model to isolate the fundamental issues in \enquote{the AI problem}. Further, so as to offer a coherent framework, we sometimes edit names of existing methods.\footnote{We recognize that others have done the same (e.g., Kopf, Zeileis, and Strobl \protect\hyperlink{ref-kopf2015framework}{2015}), and that we risk contributing to a proliferation of names.} We use a (simulated) demonstration dataset to illustrate key features of the different methods.\footnote{This data is used solely for illustrative purposes; we conduct an extensive simulation study later in the paper.} The data consists of 10,000 reference group students and 10,000 focal group students taking an eight-item test. Our simulation is similar to our multidimensional alternative to the Kopf, Zeileis, and Strobl (\protect\hyperlink{ref-kopf2015framework}{2015}) approach discussed above. Target ability is simulated as \({\theta_i}^{\text{ref}} \sim N(\mu^\text{ref} = 0,1)\) for students in the reference group and \({\theta_i}^{\text{foc}} \sim N(\mu^\text{foc} = -1,1)\) for students in the focal group. Nuisance ability is set to \({\eta_i}^{\text{ref}} = 0\) and \({\eta_i}^{\text{foc}} = -1\). The slope on target ability is set to \(a_{1j} = 1\) for all items. The slope on nuisance ability is set to \(a_{2j} = 0.5\) for the last three items (the items with DIF) and \(a_{2j} = 0\) otherwise. We can also, of course, describe these conditions from the DIF-as-item-property view where there is no nuisance ability. From that perspective, \({b_j}^{\text{foc}} = {b_j}^{\text{ref}}\) for the first five items and \({b_j}^{\text{foc}} = {b_j}^{\text{ref}} - 0.5\) for the last three items. To be sure, this data is used only for demonstrative purposes. We compare the performance of the AI methods in a more realistic scenario in the \protect\hyperlink{simstudy}{the simulation study}.

@@

Before we begin to discuss different AI methods, we begin by offering a suggestion for examining the identification problem in a given dataset. This approach is motivated by our observation that AI methods typically work like a black box. The analyst puts their item response data in and the black box outputs an identification assumption to be used when fitting subsequent models. From our perspective, tools that help an analyst to better understand their data are of great value.

To that end, we propose the \enquote{equal means, multiple imputation logit graph} (EM-MILG), which presents information about potential DIF to the analyst. EM-MILG begins by fitting a multigroup unidimensional Rasch model to the data that is identified by arbitrarily setting \(\mu^\text{foc} = 0\) (Chalmers \protect\hyperlink{ref-chalmers2012mirt}{2012}). Given that we also assume \(\mu^\text{ref} = 0\), this is equivalent to the assumption that the groups have equal mean ability. As a result, all differences in performance---either from group ability differences or DIF---manifest in the item easiness difference parameter \(\tilde{d_j}\). The tilde (e.g., \(\tilde{d_j}\)) is used to denote parameters estimated with \(\mu^\text{foc}\) set to \(0\)---as compared to the hat (e.g., \(\hat{d_j}\)), which is used for parameter estimates from a properly identified model.

@@

Figure \ref{fig:emmilg} shows the EM-MILG for the demonstration data. We emphasize that the EM-MILG contains all possible information about the difference in group performance. The analyst might assume that---because there are five items where the reference group outperforms the focal group by approximately 1 logit and only three items where the difference is 1.5 logits---items 1-5 are unbiased. This might seem to be a strong assumption, bit it is one we emphasize: All AI methods assume---in one way or another---that, if there is DIF, its a minority of items that are responsible (in this case, items 6-8). These unbiased items are known as anchor items; setting anchor items is the most common identification assumption. After anchor items have been selected, the model is refit where the identifying assumption that \(\mu^\text{foc} = 0\) is dropped and replaced with the identifying assumption that the anchor items are DIF-free (i.e., \(d_j = 0\) for anchor items). Of course, the demonstration data is designed such that there are obvious groups; this need not be the case in practice.

@@

We described a variety of AI methods and their corresponding acronyms. Some of these methods, such as AOAA and EMC, select anchor items, while others, such as MAXGI and MINBC, select an anchor point. Table \ref{table:allmethods} summarizes all of these methods. Figure \ref{fig:allmethodsfig} shows the results for each method on the demonstration data in a MILG.

@@

@@

\caption{MILG for each of the AI methods applied to the demonstration data. AOAA-OAT and EMC are identified by anchor items (represented by points). MAXGI and MINBC are identified by setting $\mu^{\text{foc}}$ set to the anchor point $\mu^{\star\text{foc}}$. AOAA and AOAA-AS are not included as they did not find an identifying assumption.}\label{fig:allmethodsfig}
\end{figure}

@@

AOAA-OAT exhibiting superior performance is an important finding given that it is, to our knowledge, not used currently. We advocate for its widespread use. One drawback of the AOAA-OAT method is that it is computationally expensive. For example, finding three items containing DIF on a 12-item test requires fitting 46 item response models, and that number grows as either the test length or the number of items testing positive for DIF grows. However, we believe the computational costs are worthwhile given the performance of this method relative to more widely implemented alternatives. To increase AOAA-OAT's use, we recommend its implementation (perhaps as the default) in popular IRT software such as the mirt R package.

@@

Moreover, future work should test these methods' performance under a greater variety of data generating conditions. For example, changing the compensatory nature of the data generating model or adding additional nuisance ability dimensions. Finally, our work focused on the Rasch model, and it is of great interest to consider how these methods extend and perform when the goal is to detect and correct for DIF when fitting other types of item response models.

@@

\leavevmode\hypertarget{ref-cook1987problems}{}%
Cook, Linda L, and Nancy S Paterson. 1987. ``Problems Related to the Use of Conventional and Item Response Theory Equating Methods in Less Than Optimal Circumstances.'' \emph{Applied Psychological Measurement} 11 (3): 225--44.

@@

\leavevmode\hypertarget{ref-furlow2009impact}{}%
Furlow, Carolyn F, Terris Raiford Ross, and Phill Gagné. 2009. ``The Impact of Multidimensionality on the Detection of Differential Bundle Functioning Using Simultaneous Item Bias Test.'' \emph{Applied Psychological Measurement} 33 (6): 441--64.

@@

\leavevmode\hypertarget{ref-kirisci2001robustness}{}%
Kirisci, Levent, Tse-chi Hsu, and Lifa Yu. 2001. ``Robustness of Item Parameter Estimation Programs to Assumptions of Unidimensionality and Normality.'' \emph{Applied Psychological Measurement} 25 (2): 146--62.

@@

\leavevmode\hypertarget{ref-topczewski2013comparison}{}%
Topczewski, Anna, Zhongmin Cui, David Woodruff, Hanwei Chen, and Yu Fang. 2013. ``A Comparison of Four Linear Equating Methods for the Common-Item Nonequivalent Groups Design Using Simulation Methods. ACT Research Report Series, 2013 (2).'' \emph{ACT, Inc.}

@@

\leavevmode\hypertarget{ref-wang2012dif}{}%
Wang, Wen-Chung, Ching-Lin Shih, and Guo-Wei Sun. 2012. ``The Dif-Free-Then-Dif Strategy for the Assessment of Differential Item Functioning.'' \emph{Educational and Psychological Measurement} 72 (4): 687--708.
